---
title: "NBA Clustering Project"
author: "Mike Kaminski"
date: '2022-05-29'
output: 
  html_document:
      code_folder: "show"
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
* library(data.table)
* library(dplyr) 
* library(ggplot2)
* library(readr)
* library(readxl)
* library(stringi)
* library(stringr)


```{r load-packages, include=FALSE}
# Libraries
library(data.table)
library(dplyr) 
library(ggplot2)
library(readr)
library(readxl)
library(stringi)
library(stringr)

```
# Import Files
* Data was scraped from [https://www.basketball-reference.com/] using Python and .csv files were saved
* Columns: blank columns need to be removed, columns containing '%' need to be renamed to 'PCT', columns containing '2' or '3' will be renamed to include 'Two' and 'Thr', and periods will be replaced by underscores

### Per 100 possessions
 * G, GS, MP, FG_PCT, ThrP_PCT, FT_PCT are not per 100 possessions, they are season totals
```{r warning=FALSE, message=FALSE}
per_poss <- read_csv("per_poss.csv")
colnames(per_poss)

per_poss1 <- subset(per_poss) %>%
  select('Year', everything()) %>%
  select(-c(contains('Unnamed'))) %>%
  rename_at(vars(contains("%")),funs(str_replace_all(., "%","_PCT"))) %>%
  rename_at(vars(contains("3")),funs(str_replace(., "3","Thr"))) %>%
  rename_at(vars(contains("2")),funs(str_replace(., "2","Two"))) %>%
  rename_at(vars(contains(" ")),funs(str_replace_all(., " ","_")))
head(per_poss1)
colnames(per_poss1)

```
### Advanced
```{r warning=FALSE, message=FALSE}
adv <- read_csv("advanced.csv")
colnames(adv)

adv1 <- subset(adv) %>%
  select('Year', everything()) %>%
  select(-c(contains('Unnamed'))) %>%
  rename_at(vars(contains("%")),funs(str_replace_all(., "%","_PCT"))) %>%
  rename_at(vars(contains("3")),funs(str_replace(., "3","Thr"))) %>%
  rename_at(vars(contains("2")),funs(str_replace(., "2","Two"))) %>%
  rename_at(vars(contains(" ")),funs(str_replace_all(., " ","_"))) 
head(adv1)
colnames(adv1)
```

### Shooting
```{r warning=FALSE, message=FALSE}
shoot <- read_csv("shooting.csv")
colnames(shoot)

shoot1 <- subset(shoot) %>%
  select(-c(contains('Unnamed'))) %>%
  select('Year', everything()) %>%
  rename_at(vars(contains(" ")),funs(str_replace_all(., " ","_"))) %>%
  rename_at(vars(contains("_of")),funs(str_replace_all(., "_of",""))) %>%
  rename_at(vars(contains("_by")),funs(str_replace_all(., "_by",""))) %>%
  rename_at(vars(contains("Distance")),funs(str_replace(., "Distance","Dis"))) %>%
  rename_at(vars(contains("-")),funs(str_replace(., "-","_"))) %>%
  rename(Avg_Dist_FGA = Dist.)

head(shoot1)
colnames(shoot1)
```

### Per Game
```{r warning=FALSE, message=FALSE}
per_game <- read_csv("per_game.csv")
colnames(per_game)

per_game1 <- subset(per_game) %>%
  select('Year', everything()) %>%
  rename_at(vars(contains("%")),funs(str_replace_all(., "%","_PCT"))) %>%
  rename_at(vars(contains("3")),funs(str_replace(., "3","Thr"))) %>%
  rename_at(vars(contains("2")),funs(str_replace(., "2","Two")))
colnames(per_game1)
head(per_game1)
```

# Exploratory Data Analysis

There's a fair amount players that we need to remove from teh dataset given their lack of contribution to the team (Ex. playing only a few games or not playing a lot of minutes).

The data includes a team name called 'TOT' which indicates that a player played or more than one team.  These players' stats appear on multiple lines - one for each team and one for the Total that season.  We need to remove the underlying teams for each player with 'TOT' as their team so we're left with only one row for each player for each season they played.  We'll create a unique index (since the index wasn't included on the scrape) based on rank (which is essentially an index value based on player name), player name and year.  We'll arrange by Games (TOT will always be the highest) and remove duplicate instances for each dataframe

### Per 100 Possesisons
```{r}
per_poss2 <- per_poss1
per_poss2$player_year <- paste(per_poss1$Rk,per_poss1$Player, per_poss1$Year) # create concatenation
per_poss2$player_year <- str_replace_all(per_poss2$player_year, fixed(" "), "") # remove spaces

per_poss3 <- per_poss2 %>% 
  arrange(Year, Player, desc(G))%>%
  filter(duplicated(player_year) == FALSE)

# compares the old df to the new df based on Team equaling 'TOT'. The TRUE valued should remain the same, the FALSE value should change because we're removing duplicates
per_poss2 %>% count(Tm == 'TOT')
per_poss3 %>% count(Tm == 'TOT')
```

### Advanced
```{r}
adv2 <- adv1
adv2$player_year <- paste(adv1$Rk,adv1$Player, adv1$Year)
adv2$player_year <- str_replace_all(adv2$player_year, fixed(" "), "")

adv3 <- adv2 %>% 
  arrange(Year, Player, desc(G))%>%
  filter(duplicated(player_year) == FALSE)

adv2 %>% count(Tm == 'TOT')

adv3 %>% count(Tm == 'TOT')

```

### Shooting
```{r}
shoot2 <- shoot1
shoot2$player_year <- paste(shoot1$Rk,shoot1$Player, shoot1$Year)
shoot2$player_year <- str_replace_all(shoot2$player_year, fixed(" "), "") 

shoot3 <- shoot2 %>% 
  arrange(Year, Player, desc(G))%>%
  filter(duplicated(player_year) == FALSE)


shoot2 %>% count(Tm == 'TOT')

shoot3 %>% count(Tm == 'TOT')
```
### Per Game
```{r}
per_game2 <- per_game1
per_game2$player_year <- paste(per_game1$Rk,per_game1$Player, per_game1$Year)
per_game2$player_year <- str_replace_all(per_game2$player_year, fixed(" "), "") 

per_game3 <- per_game2 %>% 
  arrange(Year, Player, desc(G))%>%
  filter(duplicated(player_year) == FALSE)

per_game2 %>% count(Tm == 'TOT')

per_game3 %>% count(Tm == 'TOT')
```

### Combined Dataframe

Each of the dataframes are sorted the same way for ease of processing
```{r}
per_poss3 <- per_poss3 %>% 
  arrange(Year, Player, desc(G))

adv3 <- adv3 %>% 
  arrange(Year, Player, desc(G))

shoot3 <- shoot3 %>% 
  arrange(Year, Player, desc(G))

per_game3 <- per_game3 %>% 
  arrange(Year, Player, desc(G))

```

There are duplicate columns that will need to be removed before we join the tables
```{r}
per_poss4 <- per_poss3
adv4 <- adv3
shoot4 <- shoot3
per_game4 <- per_game3
```

Year, Rank, Player, Pos, Age, TM, and G are all present in each dataframe, so those will be removed.
A suffix will be adding to each column name so that we can determine where the stat was initially located
```{r}
colnames(per_poss4)[8:32] <- paste(colnames(per_poss4)[8:32],"pp",sep="_")
colnames(adv4)[8:28] <- paste(colnames(adv4)[8:28],"adv",sep="_")
colnames(shoot4)[8:30] <- paste(colnames(shoot4)[8:30],"sho",sep="_")
colnames(per_game4)[8:31] <- paste(colnames(per_game4)[8:31],"pg",sep="_")

```

There still could be some duplicate stat columns within each of the dataframes, so apreliminary df will be created and analyzed manually to identify these instances
```{r}
dupcol_df <- per_poss4 %>%
  select(1:33) %>%
  left_join(select(adv4, c(7:29)), by = 'player_year') %>% 
  left_join(select(shoot4, c(7:31)), by = 'player_year') %>% 
  left_join(select(per_game4, c(7:32)), by = 'player_year')
dupcol_df <- dupcol_df %>% select(order(colnames(dupcol_df)))
```

##### Obvious Duplicated Columns
Columns will be removed based on a hierarchy of the dataframes: 1) Per 100, 2) Advanced, 3) Shooting, 4) Per Game

* FG_PCT: Per 100, Shooting, and Per Game
* FT_PCT: Per 100, Per Game
* G: Per 100, Advanced, Shooting, Per Game 
* GS: Per 100, Per Game
* MP: Per 100, Advanced, Shooting
* ThrP_PCT: Per 100, Per Game
* TwoP_PCT: Per 100, Per Game

```{r}
dupcol_df1 <- per_poss4 %>%
  select(1,3:33) %>% #remove rank as it's meaningless
  left_join(select(adv4, c(9:29)), by = 'player_year') %>% #remove G and MP
  left_join(select(shoot4, c(10:31)), by = 'player_year') %>% #remove G MP FG_PCT
  left_join(select(per_game4, -c(1:8,12,15,22)), by = 'player_year') #remove FG_PCT, FT_PCT, G, GS, ThrP_PCT, TwoP_PCT
colnames(dupcol_df1)
```


##### Not so Obvious Duplicated Columns
These columns are named differently, but they account for the same statistics when their values are for each player are reviewed

* ThrPAr_adv and PCT_FGA_Dis_3P_sho 
* ThrP_PCT_pp and FG_PCT_Dis_3P_sho
* TwoP_PCT_pp, FG_PCT_Dis_2P_sho,and TwoP_PCT_pg
* STL_pp and STL_PCT_adv

These duplicated columns will be removed leaving us another step closer to our final dataset
```{r}
final_df2 <- dupcol_df1 %>%
  select(-c('PCT_FGA_Dis_3P_sho','FG_PCT_Dis_3P_sho','FG_PCT_Dis_2P_sho','TwoP_PCT_pg','STL_PCT_adv'))
```

There are a ton of NA values in this dataset and it was determined that the 'less-impactful players' need to be filtered out based on games  and minutes per game.  It's likely that not all players registered every kind of stat in the games/minutes that they played, so values are appearing as NA

We'll start by finding the quartiles for games played by season and remove any player falling below the 1st quartile value, keeping only the top 75% of players each year.
```{r}
#generate quartiles values for minutes played
q <- c(0.25,0.5,0.75)
qs_games_per_year <- final_df2 %>%
  group_by(Year) %>%
  summarize(quant25 = quantile(G, probs = q[1]),
            quant50 = quantile(G, probs = q[2]),
            quant75 = quantile(G, probs = q[3])) %>%
  print(n=40)
data.frame(qs_games_per_year)

#join main df and quartile values
final_df_with_qs_g <- left_join(final_df2, qs_games_per_year[1:2], by ="Year")
```

### Plot of Games per Season
```{r}
ggplot(final_df_with_qs_g, aes(x=G, y=as.factor(Year))) + 
  geom_boxplot(color="#63727A", fill="#5A2D81", alpha=0.3) +
  labs(title="Box Plot of Games per Season",x="Games", y = "Season")+
  theme(plot.title=element_text(hjust=0.5),
        plot.subtitle=element_text(hjust=0.5)) +
  scale_x_continuous(limits = c(0,86), expand = c(0,0), breaks=seq(0, 85, 5)) 
```

The season was shorten in 2012 due to a lock-out. 2020, 2021, and 2022 were shortened due to the pandemic, so their median values are a lot different than the other seasons

Remove players that fall below the 1st quartile.  This reduces the player pool from 10,500 to 7,900
```{r}
final_df_qs_g_removed <- filter(final_df_with_qs_g, G >= quant25)
```

We'll check for NAs again.
```{r}
sapply(final_df_qs_g_removed, function(x) sum(is.na(x)))

```
These stats still have missing values:
```{r}
setDT(final_df_qs_g_removed)
names(final_df_qs_g_removed)[final_df_qs_g_removed[,sapply(.SD, function(x) any(is.na(x))),]]
```

We'll create a new dataframe to review a bit further
```{r}
final_df4 <- final_df_qs_g_removed
```

It appears that players averaging a small amount of minutes per game are still being included.  These players played in a total number of games above the 1st quartile value, but didn't really contribute much.  We'll run the same quartile method on minutes per game, using the initial final dataframe (finaldf_2) instead of the new dataframe (final_df4).  The reduction will be applied to the new dataframe.
```{r}
q <- c(0.25,0.5,0.75)
qs_mpg_per_year <- final_df2 %>%
  group_by(Year) %>%
  summarize(quant25 = quantile(MP_pg, probs = q[1]),
            quant50 = quantile(MP_pg, probs = q[2]),
            quant75 = quantile(MP_pg, probs = q[3])) %>%
  print(n=40)
data.frame(qs_mpg_per_year)

#join final_df4 and qs
final_df_qs_mpg <- left_join(final_df4, qs_mpg_per_year[1:2], by ="Year")
```
### Plot of Minutes Per Game (MPG) per Season
```{r}
boxplot(final_df_qs_mpg$MP_pg ~ final_df_qs_mpg$Year,
        main="Minutes Per Game Per Year", xlab="Year",ylab ="Minutes Per Game", horizontal=TRUE,col = "purple",las=2) 
```

Remove players that fall below the 1st quartile.  This reduces the player pool from 7,900 to 7,000
```{r}
final_df_qs_mpg_removed <- filter(final_df_qs_mpg, MP_pg >= quant25.y)
```

These stats still have missing values:
```{r}
setDT(final_df_qs_mpg_removed)
names(final_df_qs_mpg_removed)[final_df_qs_mpg_removed[,sapply(.SD, function(x) any(is.na(x))),]]
```
After review, none of these players took shots from within the above distances or in the specified type of shot.  For example, Andre Drummond shot zero three pointers during one of the years, so his ThrP_PCT_pp, PCT_FG_Astd_3P_sho, Corner_3s_PCT_3PA_sho, amd Corner_3s_PCT_3P_sho stat values were NA. The NA values were replaced with 0 for all instances.
```{r}
final_df5 <- final_df_qs_mpg_removed %>% replace(is.na(.), 0)
```
Some of the player names include an asterisk.  This asterisk will be removed and a new feature will be engineered called HOF (Hall of Fame), which is a binary column indicating that
```{r}
final_df5$HOF <- ifelse(grepl("\\*", final_df5$Player),1,0)

final_df5 <- final_df5 %>%
  mutate(Player =gsub("\\*", "", Player))
```

Some of the player names included foreign letters.  The code below removes those letters.  Additionally, periods, dashes, and single quotes were removed.
```{r}
final_df5$Player <- stri_trans_general(final_df5$Player, "Latin-ASCII")

final_df5 <- final_df5 %>%
  mutate(Player =gsub("\\.", "", Player)) %>%
  mutate(Player =gsub("\\-", "", Player)) %>%
  mutate(Player =gsub("\\'", "", Player))
```

A unique player index was created in excel, imported, and joined to the df.  The previous column that was used to remove players with more than one line per season (Tm == 'Tot') wasn't useful for identifying unique players overall.

Only valuable columns for our model were selected, columns were reordered, and a few columns were renamed.
```{r}
PlayerIndex <- read_excel("PlayerIndex.xlsx")

final_df5 <- left_join(final_df5 ,PlayerIndex,by="Player")

colnames(final_df5)
final_df5 <- final_df5 %>%
  select(c(1:31,33:88,91,103)) %>% #removes player index values and player_year column and quant25
  select('Year','Player','uniqueid', everything()) %>%
  dplyr::rename(c(Pos = Pos.x, Age = Age.x))
```

The values in the Pos (position) column sometimes included multiple positions (SG-SF, PF-C, etc.).  These values were ammended to only include the first position in the string.

```{r}
final_df5$Pos <- substr(final_df5$Pos, 1, 2)
final_df5$Pos <- str_remove_all(final_df5$Pos, "-")
final_df5 %>% distinct(Pos)
```

Final dataframe:
```{r}
write.csv(final_df5, "NBACleanData.csv")
```

