---
title: "NBA Clustering Project - Tidymodels"
author: "Mike Kaminski"
date: "2023-06-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE,
                      message = FALSE, dpi = 180,
                      fig.width = 10, fig.height = 5)
```

```{r}
library(readr)
library(dplyr) 
library(stringr)
library(ggplot2)
library(knitr)
library(tidyr)
library(stringi)
```


# Import Data

## Per 100 Possessions
```{r}
per_poss <- as.data.frame(read_csv("Data/per_poss.csv"))

per_poss1 <- per_poss %>%
  select('Year', everything(),-contains("Unnamed")) %>%
  rename_all(~ str_replace_all(., " ", "_")) %>%
  rename_all(~ str_replace_all(., "3", "Thr")) %>%
  rename_all(~ str_replace_all(., "2", "Two")) 
head(per_poss1)

```

## Advanced Stats
```{r}
adv <- as.data.frame(read_csv("Data/advanced.csv"))

adv1 <- adv %>%
  select('Year', everything(),-contains("Unnamed")) %>%
  rename_all(~ str_replace_all(., "%", "_PCT")) %>%
  rename_all(~ str_replace_all(., " ", "_")) %>%
  rename_all(~ str_replace_all(., "3", "Thr")) %>%
  rename_all(~ str_replace_all(., "2", "Two")) 
head(adv1)

```

## Shooting Stats
```{r}
shoot <- as.data.frame(read_csv("Data/shooting.csv"))

shoot1 <- shoot %>%
  select('Year', everything(),-contains("Unnamed")) %>%
  rename_all(~ str_replace_all(., " ", "_")) %>%
  rename(Avg_Dist_FGA = Dist.) %>%
  rename_all(~ str_replace_all(., "-", "_")) %>%
  rename_all(~ str_replace_all(., "_of", "")) %>%
  rename_all(~ str_replace_all(., "_by", "")) %>%
  rename_all(~ str_replace_all(., "Distance", "Dis"))
head(shoot1)
```

## Per Game
```{r}
per_game <- as.data.frame(read_csv("Data/per_game.csv"))

per_game1 <- per_game %>%
  select('Year', everything(),-contains("Unnamed")) %>%
  rename_all(~ str_replace_all(., "%", "_PCT")) %>%
  rename_all(~ str_replace_all(., "3", "Thr")) %>%
  rename_all(~ str_replace_all(., "2", "Two")) 
head(per_game1)
```

# Exploratory Data Analysis

## More Cleaning
There's a fair amount players that we need to remove from teh dataset given their lack of contribution to the team (Ex. playing only a few games or not playing a lot of minutes).

The data includes a team name called 'TOT' which indicates that a player played or more than one team.  These players' stats appear on multiple lines - one for each team and one for the Total that season.  We need to remove the underlying teams for each player with 'TOT' as their team so we're left with only one row for each player for each season they played.  We'll create a unique index (since the index wasn't included on the scrape) based on rank (which is essentially an index value based on player name), player name and year.  We'll arrange by Games (TOT will always be the highest) and remove duplicate instances for each dataframe

### Per 100 Possessions
```{r}
per_poss2 <- per_poss1 %>%
  mutate(player_year = paste0(Rk,Player,Year)) %>% # creates an ID
  mutate(player_year = str_replace_all(player_year," ","")) #removes spaces

per_poss3 <- per_poss2 %>% 
  arrange(Year, Player, desc(G))%>%
  filter(duplicated(player_year) == FALSE)

# compares the old df to the new df based on Team equaling 'TOT'. The TRUE valued should remain the same, the FALSE value should change because we're removing duplicates
per_poss2 %>% count(Tm == 'TOT')
per_poss3 %>% count(Tm == 'TOT')

```

### Advanced
```{r}
adv2 <- adv1 %>%
  mutate(player_year = paste0(Rk,Player,Year)) %>% # creates an ID
  mutate(player_year = str_replace_all(player_year," ","")) #removes spaces

adv3 <- adv2 %>% 
  arrange(Year, Player, desc(G))%>%
  filter(duplicated(player_year) == FALSE)

# compares the old df to the new df based on Team equaling 'TOT'. The TRUE valued should remain the same, the FALSE value should change because we're removing duplicates
adv2 %>% count(Tm == 'TOT')
adv3 %>% count(Tm == 'TOT')
```

### Shooting
```{r}
shoot2 <- shoot1 %>%
  mutate(player_year = paste0(Rk,Player,Year)) %>% # creates an ID
  mutate(player_year = str_replace_all(player_year," ","")) #removes spaces

shoot3 <- shoot2 %>% 
  arrange(Year, Player, desc(G))%>%
  filter(duplicated(player_year) == FALSE)

# compares the old df to the new df based on Team equaling 'TOT'. The TRUE valued should remain the same, the FALSE value should change because we're removing duplicates
shoot2 %>% count(Tm == 'TOT')
shoot3 %>% count(Tm == 'TOT')
```

### Per Game
```{r}
per_game2 <- per_game1 %>%
  mutate(player_year = paste0(Rk,Player,Year)) %>% # creates an ID
  mutate(player_year = str_replace_all(player_year," ","")) #removes spaces

per_game3 <- per_game2 %>% 
  arrange(Year, Player, desc(G))%>%
  filter(duplicated(player_year) == FALSE)

# compares the old df to the new df based on Team equaling 'TOT'. The TRUE valued should remain the same, the FALSE value should change because we're removing duplicates
per_game2 %>% count(Tm == 'TOT')
per_game3 %>% count(Tm == 'TOT')
```

There are duplicate column names in each df, so need to change names to include a unique identifier
```{r}
adv4 <- adv3 %>%
  rename_at(vars(8:28), ~paste0(.,"_adv"))

per_game4 <- per_game3 %>%
  rename_at(vars(9:31), ~paste0(.,"_pg"))

per_poss4 <- per_poss3 %>%
  rename_at(vars(9:32), ~paste0(.,"_pp"))

shoot4 <- shoot3 %>%
  rename_at(vars(8:29), ~paste0(.,"_sho"))

```

There still could be some duplicate stat columns within each of the dataframes, so a preliminary df will be created and analyzed manually to identify these instances
```{r}
dupcol_df <- per_poss4 %>%
  select(1:33) %>%
  left_join(select(adv4, c(7:29)), by = 'player_year') %>% 
  left_join(select(shoot4, c(7:31)), by = 'player_year') %>% 
  left_join(select(per_game4, c(7:32)), by = 'player_year')
dupcol_df <- dupcol_df %>% select(order(colnames(dupcol_df)))
```

#### Obvious Duplicated Columns
Columns will be removed based on a hierarchy of the dataframes: 1) Per 100, 2) Advanced, 3) Shooting, 4) Per Game

* FG_PCT: Per 100, Shooting, and Per Game
* FT_PCT: Per 100, Per Game
* G: Per 100, Advanced, Shooting, Per Game 
* GS: Per 100, Per Game
* MP: Per 100, Advanced, Shooting
* ThrP_PCT: Per 100, Per Game
* TwoP_PCT: Per 100, Per Game

```{r}
dupcol_df1 <- per_poss4 %>%
  select(1,3:33) %>% #remove rank as it's meaningless
  left_join(select(adv4, c(9:29)), by = 'player_year') %>% #remove G and MP
  left_join(select(shoot4, c(10:31)), by = 'player_year') %>% #remove G MP FG_PCT
  left_join(select(per_game4, -c(1:8,12,15,22)), by = 'player_year') #remove FG_PCT, FT_PCT, G, GS, ThrP_PCT, TwoP_PCT
```


#### Not so Obvious Duplicated Columns
These columns are named differently, but they account for the same statistics when their values are for each player are reviewed

* ThrPAr_adv and PCT_FGA_Dis_3P_sho 
* ThrP_PCT_pp and FG_PCT_Dis_3P_sho
* TwoP_PCT_pp, FG_PCT_Dis_2P_sho,and TwoP_PCT_pg
* STL_pp and STL_PCT_adv

These duplicated columns will be removed leaving us another step closer to our final dataset
```{r}
final_df2 <- dupcol_df1 %>%
  select(-c('PCT_FGA_Dis_3P_sho','FG_PCT_Dis_3P_sho','FG_PCT_Dis_2P_sho','TwoP_PCT_pg','STL_PCT_adv')) %>%
  select(player_year, everything())
```

It's likely that players who didn't play a ton of games will impact the analysis, so it would be beneficial to remove these players.

```{r}
final_df2 %>%
  ggplot(aes(Year, G)) +
  geom_point()
```

I'll start by looking at the number of games played each year.  Some season were shortened due to the lockout or to covid, so picking an arbitrary number might not be the best idea.
```{r}
q <- c(0.25,0.5,0.75)
qs_games_per_year <- final_df2 %>%
  group_by(Year) %>%
  summarize(min = min(G),
            quant25 = quantile(G, probs = q[1]),
            quant50 = quantile(G, probs = q[2]),
            quant75 = quantile(G, probs = q[3]),
            max = max(G)) %>%
  print(n=40)
# kable(qs_games_per_year)

```

Players could have played more than 82 games in a year due to trades mid-season.

```{r}
#creates a new df
final_df_with_qs_g <- left_join(final_df2, qs_games_per_year[1:4], by ="Year")
```

```{r}
ggplot(final_df_with_qs_g, aes(x=G, y=as.factor(Year))) + 
  geom_boxplot(color="#63727A", fill="#5A2D81", alpha=0.3) +
  labs(title="Box Plot of Games per Season",x="Games", y = "Season")+
  theme(plot.title=element_text(hjust=0.5),
        plot.subtitle=element_text(hjust=0.5)) +
  scale_x_continuous(limits = c(0,86), expand = c(0,0), breaks=seq(0, 85, 5)) 
```

### Create a cutoff
Removing players at the 25% percentile or 50% percentile would likely make the most sense.
```{r}
final_df_with_qs_g %>%
  filter(G >= quant25) %>%
  summarize(count = n())
# 7939

final_df_with_qs_g %>%
  filter(G >= quant50) %>%
  summarize(count = n())
# 5362  

# It's likely the NA values are because a player didn't have any stats for that, so I will replace all NAs with 0
final_df_with_qs_g %>%
  filter(G >= quant25) %>%
  summarise(sum(is.na(.)))

final_df_with_qs_g %>%
  filter(G >= quant50) %>%
  summarise(sum(is.na(.)))
  
final_df_with_qs_g <- final_df_with_qs_g %>%
  mutate_all(~ replace_na(., 0))

```

Given the data and the obejctive, it's likely better to use a minutes cutoff rather than a games played cutoff.  This will better account for good players that got injured that played fewer games. 12 minutes will be the cutoff and can always be changed to inlcude a 20 mpg cutoff.


I'll look at minutes as well.
```{r}
q <- c(0.25,0.5,0.75)
qs_mins_per_year <- final_df2 %>%
  group_by(Year) %>%
  summarize(min = min(MP_pg),
            quant25 = quantile(MP_pg, probs = q[1]),
            quant50 = quantile(MP_pg, probs = q[2]),
            quant75 = quantile(MP_pg, probs = q[3]),
            max = max(MP_pg)) %>%
  print(n=40)
```

```{r}
# Above 12 mins per game or above 20 mins per game might be an appropriate cutoff.  12 mins is one quarter
mean(qs_mins_per_year$quant25) #12.2
mean(qs_mins_per_year$quant50) #19.7
```

```{r}
#creates a new df
final_df_with_qs_mpg <- left_join(final_df2, qs_mins_per_year[1:4], by ="Year")
```

```{r}
final_df_with_qs_mpg %>%
  filter(MP_pg >= 12) %>%
 summarize(count = n())
# 7989

final_df_with_qs_mpg %>%
  filter(MP_pg >= 20) %>%
  summarize(count = n())
# 5171 

# It's likely the NA values are because a player didn't have any stats for that, so I will replace all NAs with 0
final_df_with_qs_mpg %>%
  filter(MP_pg >= 12) %>%
  summarise(sum(is.na(.)))

final_df_with_qs_mpg %>%
  filter(MP_pg >= 20) %>%
  summarise(sum(is.na(.)))
  
final_df_with_qs_mpg <- final_df_with_qs_mpg %>%
  mutate_all(~ replace_na(., 0))
```

```{r}
cutoff_df <- final_df_with_qs_mpg %>%
  select(-min,-quant25,-quant50) %>% #removes unneeded variables
  filter(MP_pg >= 12) %>%
  mutate(HOF = ifelse(grepl("\\*", Player),1,0)) %>% #creates a new variable called Hall of Fame
  mutate(Player = stri_trans_general(Player, "Latin-ASCII")) %>%
  mutate(Player = gsub("[.'*]", "", Player)) %>%
  mutate(Player = gsub("\\-", " ", Player)) 

a <- cutoff_df %>% distinct(Player)

```





Data was taken from the previous doc
```{r}
library(readr)
NBACleanData <- read_csv("Data/NBACleanData.csv",show_col_types = FALSE)
  
```


```{r}
nba <- NBACleanData %>%
  select(-c(1:10,"HOF","PF_pg","PF_pp","Heave_Att_sho","Heaves_Comp_sho","Dunks_PCT_FGA_sho","Dunks_sho")) #remove unwanted variables
# colnames(nba)

nba_2022 <- NBACleanData %>%
  filter(Year == 2022) %>%
  select(-c(1:10,"HOF","PF_pg","PF_pp","Heave_Att_sho","Heaves_Comp_sho","Dunks_PCT_FGA_sho","Dunks_sho")) #remove unwanted variables
  

```


```{r}
# trial for removing variables are seem to be duplicative
nba <- NBACleanData %>%
  select(-c(1:10,"HOF","PF_pg","PF_pp","Heave_Att_sho","Heaves_Comp_sho","Dunks_PCT_FGA_sho","Dunks_sho")) %>% #remove unwanted variables 
  select(
    -FG_pp, #field goals made per 100 possessions
    -ThrP_pp, # three pointers made
    -TwoP_pp, # two pointers made
    -FT_pp, # free throws made
    -contains("_pg"), # removes per game
    -contains("FG_PCT"),
    -Avg_Dist_FGA_sho,
    -contains("Corner"),
    -contains("PCT_FG_Ast"),
    -TRB_pp
  ) 

```

```{r}
normalize <- function(x)
{
  return((x- min(x)) /(max(x)-min(x)))
}
nba_normal <- as.data.frame(lapply(nba, FUN=normalize))
# nba_normal_2022 <- as.data.frame(lapply(nba_2022, FUN=normalize))
```



```{r}
library(tidymodels)

set.seed(525)

kclust <- kmeans(nba_normal, centers = 4)

# kclust_2022 <- kmeans(nba_normal_2022, centers = 7)

```

```{r}
library(broom)
tidy(kclust)
# tidy(kclust_2022)
```


```{r}
augment(kclust, nba_normal) %>%
  ggplot(aes(FGA_pp, TwoPA_pp, color = .cluster)) +
  geom_point()

# augment(kclust_2022, nba_normal_2022) %>%
#   ggplot(aes(PTS_pg, VORP_adv, color = .cluster)) +
#   geom_point()
```


```{r}
augment(kclust, nba_normal)
tidy(kclust)
glance(kclust)


# augment(kclust_2022, nba_normal_2022)
# tidy(kclust_2022)
# glance(kclust_2022)
```

```{r}
kclusts <- 
  tibble(k = 1:15) %>%
  mutate(
    kclust = map(k, ~ kmeans(nba_normal, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, nba_normal)
  )

kclusts

# kclusts_2022 <- 
#   tibble(k = 1:15) %>%
#   mutate(
#     kclust_2022 = map(k, ~ kmeans(nba_normal_2022, .x)),
#     tidied = map(kclust_2022, tidy),
#     glanced = map(kclust_2022, glance),
#     augmented = map(kclust_2022, augment, nba_normal_2022)
#   )
# 
# kclusts_2022
```


```{r}
kclusts %>%
  unnest(glanced) %>%
  ggplot(aes(k, tot.withinss)) +
  geom_line(alpha = 0.8) +
  geom_point(size = 2)

kclusts %>%
  unnest(glanced) %>%
  ggplot(aes(k, tot.withinss-lead(tot.withinss))) +
  geom_line(alpha = 0.8) +
  geom_point(size = 2)

kclusts %>%
  unnest(glanced) %>%
  ggplot(aes(k, tot.withinss-lead(tot.withinss,n=2))) +
  geom_line(alpha = 0.8) +
  geom_point(size = 2)


kclusts %>%
  unnest(glanced) %>%
  ggplot(aes(k, tot.withinss)) +
  geom_line(alpha = 0.8) +
  geom_point(size = 2) +

  geom_line(aes(k, tot.withinss-lead(tot.withinss)), alpha = 0.8 ) +
  geom_point(aes(k, tot.withinss-lead(tot.withinss)),size = 2,color = "red") +

  geom_line(aes(k, tot.withinss-lead(tot.withinss,n=2)), alpha = 0.8) +
  geom_point(aes(k, tot.withinss-lead(tot.withinss,n=2)),size = 2, color = "blue")
```



```{r}
clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))

clusters_2022 <- 
  kclusts_2022 %>%
  unnest(cols = c(tidied))

assignments_2022 <- 
  kclusts_2022 %>% 
  unnest(cols = c(augmented))

clusterings_2022 <- 
  kclusts_2022 %>%
  unnest(cols = c(glanced))


```

```{r}
p1 <- 
  ggplot(assignments,aes(x=FG_pp, y = FGA_pp)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```

# hierarchical
```{r}
library(workflows)
library(parsnip)
library(tidyclust)
library(tidyverse)
```

```{r}

```

